---
title: 'Photosynthesis and Cellular Respiration'
author: ''
date: '2021-01-24'
slug: photosynthesis
categories: ['VCE', 'Biology', 'Unit 3', 'Area of study 3']
tags: ['High school','VCE', 'Photosynthesis', 'Cellular Respiration', 'Tutoring']
subtitle: ''
summary: "Let's focus on what you actually need to know on everyone's favourite VCE Biology topic (sarcasm btw)"
authors: []
lastmod: '2021-01-24T04:14:47+11:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

<p style="font-size">

Note: This page is adapted from a learning plan + notes I used to teach online VCE summer school 2021. It's now been (mostly) adapted for the 2022-2026 VCAA study design. 
</p>


<h2 style="color:#D32F2F"> Learning Objectives </h2>

*By the end of this lesson you should be able to:* 

- Contrast the purpose of aerobic respiration, anaerobic respiration and photosynthesis
- List the main inputs and outputs for the light dependent and light independent stages of photosynthesis
- List the main inputs, outputs and ATP yields for stages of cellular respiration *(glycolysis, anaerobic respiration, Kreb's cycle, electron transport chain in yeast and in animals)*
- Explain how evidence ****(including chromosomal structure, double membrane, ribosome type) supports the bacterial origins of mitochondria and chloroplasts

- (Bonus Easy) Agree it's ridiculous that life exists given the complexity of reactions it needs (photosynthesis, cellular respiration)
- (Bonus Challenge) Explain how different factors (temperature,  light, glucose availability, carbon dioxide and oxygen concentration) impact the rate of cellular respiration and photosynthesis

<br>
<br>

<h2 style="color:#D32F2F">
Relevant VCAA 2022 Study Design Dot Points:    
</h2> 

- *inputs, outputs and locations* of the light dependent and light independent stages of photosynthesis in C3 plants (details of biochemical pathway mechanisms are not required) 
- the *role of Rubisco in photosynthesis*, including adaptations of C3, C4 and CAM plants to maximise the efficiency of photosynthesis
- the *factors that affect the rate of photosynthesis*: light availability, water availability, temperature and carbon dioxide concentration
- the *main inputs, outputs and locations* of glycolysis, Krebs Cycle and electron transport chain including ATP yield (details of biochemical pathway mechanisms are not required)
- the *location, inputs and the difference in outputs* of anaerobic fermentation in animals and yeasts 
- the *factors that affect the rate of cellular respiration*: temperature, glucose availability and oxygen concentration

... this looks like a lot to know - but don't worry, it's actually not at all - it's just boring.           

<br>

<h3> Important VCAA Note </h3>

> VCAA emphasises that ***"details of the biochemical pathway mechanisms are not required"

What this means that what you really need to know is: WHERE, WHAT GOES IN, WHAT COMES OUT and HOW MUCH ATP for each stage *(how much ATP is only for cellular respiration)*. Sooo ... let's just focus on these aspects firstly. 

<br>
<br>
<br>

<h2 style="color:#D32F2F">
🗣️ Cellular Respiration
</h2>

[Crash course video: ATP and Respiration ](https://www.youtube.com/watch?v=00jbG_cfGuQ&vl=en)

<h3>
Cellular respiration equation:
</h3> 

Chemical equation:

$$ 6O_{2} + C_{6}H_{12}O_{6} + 36ADP + 36P_{i} → 6CO_{2} + 6H_{2}O + 36ATP $$
Word equation:
$$ \text{Oxygen} + \text{Glucose} + \text{ADP} + \text{Inorganic Phosphorus} → \text{Carbon Dixode} + \text{Water} + \text{ATP (avaliable energy)} $$ 
<h3> Most important table to remember for cellular respiration </h3>

```{r, include = FALSE, echo = FALSE}

library(readr)
library(dplyr)
library(kableExtra)
what_to_know_for_cellular_respiration <- read_csv("../../../static/media/tables/what-to-know-for-cellular-respiration.csv")
```

```{r, echo = FALSE}

what_to_know_for_cellular_respiration %>% 
  knitr::kable(escape = FALSE) %>% 
  scroll_box(width = "100%", height = "100%")

```


<br>
<br>
<br>

<h2 style="color:#D32F2F"> 🌻 Photosynthesis </h2>

TBA ...

<h2 style="color:#D32F2F"> Extra Resources </h2>

- **Douchy's VCE Biology Podcast**
  - [Photosynthesis](https://open.spotify.com/episode/4lWbTSpKtmtzBLykDvgpUD?si=ucBi_S9TSTiNHMZC0mFJmg)
  - [Cellular respiration](https://open.spotify.com/episode/4J1PVX5P6Kcw0C1wXSnYKI?si=0KUPwY52RcW26B2vporQuA)

- [Energy transformations notes](https://freevcenotes.com/biology/notes/energy-transformations)